\LARGE \bf Enhancing Security in Multi-Robot Systems through Co-Observation Planning on Unsecured Trajectories

### 2023-09-11 paper.tex
update

\begin{figure}
	\centering
    \subfloat[Co-observation \label{fig:example-co-observation}]{\includegraphics[width=0.45\linewidth]{co-observ}}
    \subfloat[Cross-trajectory co-observation \label{fig:example-cross-traj}]{\includegraphics[width=0.45\linewidth]{cross-traj}}

@@

To give a concrete example of the cost $\varPhi$ and the set $\Omega$, we introduce a representative application that will be used for all the simulations throughout the paper.
\begin{example} Assume that there exists a slowly-time-varying scalar or vector field $\vx$ whose value needs to be estimated at a given number of discrete locations. To explore unknown spatially-distributed fields, autonomous agents are usually required to traverse the entire unknown environment while collecting sensory data to estimate a corresponding map (Figure~\ref{fig:ADMM-example-application}). We model the map as a set of locations of interests arranged on a regular grid, where at each location, we assume that there is a corresponding slowly-time-varying quantity. For our purposes, we are interested in finding paths that best reconstruct the field, i.e., that achieve the minimum uncertainty. 
We address the estimation problem using Kalman Filters (KFs)(cf. \cite{anderson2012optimal}), with a focus on the estimated covariance $P_j$, to calculate the map model uncertainty. We assume the robot provides field measurements at waypoints~$q$, with Gaussian radial basis function modeling radius and quality.
\rtron{The previous paragraph can be written more compactly.}

@@

  \begin{definition} Let $\nu_\cF$ and $\nu_\cE$ be two unitary vectors ($\norm{\nu_\cF}=\norm{\nu_\cE}=1$). Define the normalized vector $u$ as
    \begin{equation}
      u'=\nu_\cF+\nu_\cE,\quad u =\frac{u'}{\norm{u'}}.
    \end{equation}
    The \emph{Householder rotation} $H(\nu_\cF,\nu_\cE)$ is defined as
    \begin{equation}\label{eq:H definition}
      H(\nu_\cF,\nu_\cE) = 2 u u\transpose-I.
    \end{equation}
  \end{definition}
  The main property of interest for our application is the fact that $H$ is a rotation mapping $\hat{\nu_\cF}$ to $\hat{\nu_\cE}$, as shown by the following.
  \begin{proposition}\label{prop:HProperidy}
    The matrix $H$ has the following properties:
    \begin{enumerate}
    \item It is a rotation, i.e.
      \begin{enumerate}
      \item\label{it:orthonormality} $H\transpose H=I$;
      \item\label{it:determinant} $\det(H)=1$.
      \end{enumerate}
    \item\label{it:transformation} $\nu_\cE=H \nu_\cF$.
    \end{enumerate}
  \end{proposition}

  We compute the differential of $H$ implicitly using the relation \eqref{equ:dt_to_dx}. We will use the notation $\cross{v}:\real{3}\to\real{3\times 3}$ to denote the matrix representation of the cross product with the vector $v$, i.e.,
  \begin{equation}
    \bmat{v_1\\v_2\\v_3}\mapsto \bmat{0 & -v_3 &v_2\\v_3 & 0 & -v_1\\-v_2 & v_1 & 0},
  \end{equation}
  such that $[v]_\times w=v\times w$ for any $w\in\real{3}$. %One can verify by direct computation the following property:
  %\begin{equation}\label{eq:asymmetric to cross}
  %  wv\transpose-vw\transpose=\cross{\cross{v}w}.
  %\end{equation}
  \begin{proposition}\label{prop:Hderivitive}
    Let $\nu_\cF(t)$ represent a parametric curve. Then we have
    \begin{equation}
      \dot{H}=H\cross{-2M\dot{\nu}_{\cF}}
    \end{equation}
    where the matrix $M\in\real{3\times3}$ is given by
    \begin{equation}
      M=[u]_{\times}  \frac{ \left( I - u u\transpose \right) \left( I- \nu_\cF \nu_\cF\transpose \right)} {\norm{u'} \norm{\nu_\cF}}.
    \end{equation}
  \end{proposition}
\rtron{State where the proof can be found}

@@

We now provide the functions $D(\vq)$, the sets $\cZ$, and the corresponding projection operators $\Pi_\cZ$ for traditional path planning constraints (\Cref{sec:??}-\ref{sec:??}), co-observation security constraints (\Cref{sec:??}), and reachability constraints (\Cref{sec:??}). The latter are based on the definition of \emph{ellipsoidal reacability regions} (\Cref{sec:??}).
\subsection{Velocity constraints}
  The movement of each agent is constrained by a maximum distance in any direction over a single discrete time step. We then define the function $D(\vq)$ to return the velocity vectors for the $i$-th agent at time step $j$; its result then needs to be constrained to a sphere.

  \begin{constraint}[Velocity constraint]
\begin{align}
     D_{ij}(\vq) &= q_{ij}-q_{i(j-1)}, \quad j\in\{1,\ldots, T\}\\
\label{eq:velocity_constraint}
   \sZ_{ij} &=\bigl\{z \mid \norm{z} \leq v_{max}\bigr\}\\
   \Pi_{\sZ_{ij}}(z)&=\begin{cases}
   v_{\max}\frac{z}{\norm{z}} \textrm{ if } \norm{z}>v_{\max}\\
   z\textrm{ otherwise}
   \end{cases}
\end{align}
\end{constraint}
\rtron{Put all the constraints using this \LaTeX environment.}
\rtron{Make sure notation for $\sZ$ is consistent}
%The projection operator $\Pi_\zeta(z)$ for this constraint  is simply a projection of $z$ to the interior of a sphere centered at the origin and having radius $v_{max}$. This constraint could be adapted to enforce more refined dynamical models.

%implies that the projection of the vector $z$ is inside a sphere with a radius of $v_{max}$ which can then be written as:
%\begin{equation}\label{eq:velocity_constraint_update}
%    \Pi_{\zeta}(z) = \begin{cases}
% v_{max}\frac{z}{\norm{z}} &\text{if } \norm{z} > v_{max},\\
%z	& otherwise.
%\end{cases}
%\end{equation}

\subsection{Convex polygonal obstacles collision constraints}

@@

\begin{constraint}[Obstacle constraint]
\begin{align}
     	D(q_{ij}) &= \max_{k=1,\dots,l} (d_{k}(q_{ij},\cP))\\
\label{eq:obstacle_constraint}
  \sZ &= \{z \mid z \geq 0 \},\\
   \Pi_\sZ(z) & = \begin{cases}
   0 & \textrm{if} \quad z < 0\\
   z  & \textrm{otherwise}
   \end{cases}
\end{align}
\end{constraint}

The motivating idea for \eqref{eq:zoneConstraintMap} and \eqref{eq:zoneConstraintProjection} is to identify waypoints within a specific region and project them to the closest boundary. Non-convex obstacles can be handled by the union of (possibly overlapping) convex obstacles. By setting the normal vector pointing toward the inside of the region, this method can also serve to constrain the robots within the designated workspace region.

\subsection{Waypoints with flexible deadlines}
It is common for security related constraints to require robot reach a specific location at specific time. We assume robot $i$ is required to reach a given point $p$ within a radius of $d_{max}$ at some time instant $j$ belonging to a given time window $[t_1,t_2]$. The constraint function $D(\vq)$ to return the smallest distance from the point $p$ to waypoint on the trajectory restricted to the relevant time window. 
\begin{constraint}[Waypoint constraint]
\begin{align}
     	D(\vq) &= \min_{j\in\{t_1,\dots,t_2-1\}}\bigl(\dist(p,\overrightarrow{q_{ij}q_{i(j+1)}})\bigr)\\
\label{eq:obstacle_constraint}
  \sZ &= \{z \mid z < z_{max} \},\\
   \Pi_\sZ(z) & = \min(z, d_{max}).
\end{align}
\end{constraint}

where $\dist(p,\overrightarrow{q_{ij}q_{i(j+1)}})$ returns the distance between the fixed point $p$ and the segment $\overrightarrow{q_{ij}q_{i(j+1)}}$. Note that this function returns the smallest distance between $(p,q_{ij})$ and $(p,q_{i(j+1)})$ if the projection of the point $p$ does not lie on the line segment~$\overrightarrow{q_{ij}q_{i(j+1)}}$; as a consequence, this constraint does not need to be satisfied exactly at one of the points on the discretized trajectory, but it can also be satisfied ``en route'' on the segment between them.

%Note that \eqref{eq:waypoint_constraint} and \eqref{eq:waypoint_constraint_update} are equivalent to \eqref{eq:velocity_constraint} and \eqref{eq:velocity_constraint_update}, except for the fact that the quantity in \eqref{eq:waypoint_constraint_D} is always a positive scalar.

\rtron{Give example application here, and include application of ADMM without the security constraints, but highlighting a path that would break the security}
\zyang{plot new simpler example, add collision avoidance}

%\subsection{Secure Multi-robot path planning using co-observation and reachability constraints}
\subsection{Co-observation schedule constraint}
The co-observation constraints is used to guarantee that, at time required by the schedule, two robots should get close enough with each other to observe each other's behavior. This constraint is modeled as a relative distance constraint between two robots at some time instant $j$ (i.e., to require two agents see each other within a certain radius to inspect each other, or to exchange data). The location of the co-observation location should make sure that the ellipsoidal reachability region between each co-observation, considered in later sections \ref{sec:reachability}, has an empty intersection with all forbidden regions.
 We write the function for the constraint as:
\rtron{Use the constraint environment}
\begin{align}
D(\vq) = \overrightarrow{q_{aj}q_{bj}},
\end{align}
where $a,b$ are the indices of the pair of agents required for a mutual inspection. The corresponding constraint set is simply
\begin{equation}
\sZ= \{z \mid z \leq d_{max} \},
\end{equation}
with a projection function:
\begin{equation}\label{eq:co-observation_constraint_update}
\Pi_{\sZ}(z) = \begin{cases}
d_{max}\frac{z}{\norm{z}} &\text{if } \norm{z} > d_{max},\\
z	& otherwise.
\end{cases}
\end{equation}
The locations $q_{aj}$ and $q_{bj}$ where the co-observation performed are computed as part of the optimization.

@@

The reachability ellipsoid is an over-approximation of the exact reachability region; the difference between the two is due to the discretization of the trajectory, and the fact that $\cE$ does not consider the presence of obstacles.
\subsubsection{Transformation to canonical coordinates}\label{sec:rotation2Standard}
  To simplify the problem, a canonical rigid body transformation is used to transform the ellipse $\cE$ from a global frame $\cF$ to a canonical frame $\cF_\cE$. The latter is defined such that the center of the ellipsoid is located at the origin and the foci are aligned with the first axis of $\cF_\cE$. Since the transformation depends on the two foci, i.e., two waypoints of an agent, the challenge here is to construct the transformation in a differentiable way.
  For the convenience of derivation, we define the coordinate transformation from $\cF$ to $\cF_\cE$ using a rotation $R^\cF_\cE$ and a translation $o^\cF_\cE$, which, to simplify the notation, from now on we simply refer to as $R$ and $o$, respectively. The transformation of a point from the frame $\cF_\cE$ to the frame $\cF$ and its inverse are given by the formulation:
  \begin{equation}
    \Fframe{q}=R\Eframe{q}+o,\label{eq:E to F transform}\quad
    \Eframe{q}=R\transpose(\Fframe{q}-o).
  \end{equation}
 % The reverse transformation is $\Eframe{q}=R\transpose( \Fframe{q}-o)$.

  We define $\nu_\cF$ and $\nu_\cE$ to represent the $x$-axis unitary vector of $\cF_\cE$ in the frames $\cF$ and $\cF_\cE$, respectively. Formally:
  \begin{align}
    \nu'_\cF&=q_2-q_1, & \nu_\cF &= \frac{\nu'_\cF}{\norm{\nu_\cF}}, & \nu_\cE&=[1,0,0]^T;
  \end{align}
  see Fig.\ref{fig:EllipseConstraintExample} for an illustration. Note that $\nu_\cE$ is constant while, for the sake of clarity, $\nu_{\cF}$ depends on $q_1,q_2$. We then define the rotation $R$ using a Householder rotation, while from \eqref{eq:E to F transform} we see that $o$ represents the center of $\cE(q_1,q_2)$ expressed in $\cF$, i.e.:
  \begin{align} 
    R&=H(\nu_\cF(q_1,q_2),\nu_\cE(q_1,q_2)), &o&=\frac{1}{2}(q_1+q_2).
  \end{align}
  To simplify the notation, in the following we will consider $H$ to be a function of $q_1,q_2$ directly, i.e. $H(q_1,q_2)$. 
  The ellipse $\cE$ expressed in $\cF_\cE$ is given by $\Eframe{\cE}=\{\Eframe{q}\in\real{m}:d(q^\cE_1,\Eframe{q})+d(\Eframe{q},q^\cE_2)<2a\}$, where the coordinates of the two foci $q^\cE_1,q^\cE_2$ in $\cF_\cE$ are
  \begin{align}
    q^\cE_1=\bmat{c & 0 & 0}\transpose, && q^\cE_2=\bmat{-c & 0 & 0}\transpose,
  \end{align}
  with $c=\frac{\norm{q_2-q_1}}{2}$.

\rtron{You can use a definition to mirror definition 3}
  The ellipsoid $\cE$ in the canonical frame can be described as the zero level set of the quadratic function
  \begin{equation}\label{equ:standard-ellipse}
    \Eframe{E}(\Eframe{q}) = \Eframe{q}\transpose Q \Eframe{q} - 1
  \end{equation}
  where
  \begin{equation}
    Q = \diag(a^{-2},b^{-2},b^{-2}),
  \end{equation}
  and $b = \sqrt{a^2-c^2}$.
  The ellipse parameters $a$,$b$ represent the lengths of the major axes.

@@

\subsection{Point-ellipsoid reachability constraint}\label{sec:ellipsoid-point}
As shown in Fig.\ref{fig:Ellipse-to-point}, we consider a forbidden region in the shape of a single point $q_{avoid}$. The goal is to design the trajectory $q(t)$ such that $q_{avoid}\notin\mathcal{E}(q_1,q_2)$. We first define a projection function $\pi_{p\cE}(q_{avoid};q_1,q_2,a) = q_p$, which returns a projected point $q_p$ of $q_{avoid}$ outside the ellipse, i.e., as the solution to

\rtron{For these constraints, add $\pi$ inside the constraint environment, then follow that environment with the proposition with the differential of $\pi$.}

\begin{equation}\label{eq:ellipse-point-projection}
\pi_{p\cE}=\argmin_{q_p\in\cE^c} \quad \norm{q_{avoid}-q_p}^2 
\end{equation}
where $\cE^{c}$ is the set complement of the region $\cE$.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\linewidth]{Ellipse2point}
\caption{Point-ellipsoid constraint, a point $f$ inside ellipsoid is projected to the areas outside the ellipsoid $f_{p}$.}
\label{fig:Ellipse-to-point}
\end{center}
\end{figure}

  Then, the constraint $q_{avoid}\notin\mathcal{E}(q(t_1),q(t_2))$ can be written as:
  \begin{equation}
    D(q) = \pi_{p\cE}(q_{avoid};q(t_1),q(t_2),r)-q_{avoid} = 0
  \end{equation}

  For cases where $q_{avoid} \notin \cE(q(t_2),q(t_1),r)$, $\pi_{p\cE}(q_{avoid}) = q_{avoid}$. And for cases where $q_{avoid} \in \cE(q_1,q_2,r)$, $D(q)$ needs to be projected to the boundary of the ellipse, which is discussed below.

@@

  The point to project, $q_{avoid}$, can be likewise expressed in $\cF_\cE$ as $\Eframe{q}_{avoid} =  H (\Fframe q_{avoid}-o)$.

  We now turn our attention to the problem of projecting $\Eframe{q}_{avoid}$ on the zero level set of $E^\cE$ (i.e., the reachability ellipsoid in the canonical frame). The derivations below are loosely inspired by \cite{eberly}.

>>>>

\begin{figure}
	\centering
    \subfloat[Plan-deviation attack\label{fig:example-plan-dev}]{\includegraphics[width = 0.32\linewidth]{plan-deviation}}
    \subfloat[Co-observation \label{fig:example-co-observation}]{\includegraphics[width=0.32\linewidth]{co-observ}}
    \subfloat[Cross-trajectory co-observation \label{fig:example-cross-traj}]{\includegraphics[width=0.32\linewidth]{cross-traj}}

@@

To give a concrete example of the cost $\varPhi$ and the set $\Omega$, we introduce a representative application that will be used for all the simulations throughout the paper.

\begin{example} 
We consider the estimation of a slowly-varying scalar or vector field, denoted as $\vx$, at discrete locations within an unknown environment. Autonomous agents navigate this environment, collecting sensory data to construct a corresponding map (see Figure~\ref{fig:ADMM-example-application}). Our map comprises points of interest arranged on a grid, each associated with a slowly-changing value. Our goal is to find paths that minimize uncertainty and effectively reconstruct the field. We utilize Kalman Filters (KFs)~(cf. \cite{anderson2012optimal}) to estimate uncertainty through the covariance $P_j$. Additionally, the robot provides field measurements at waypoints $q$ using a Gaussian radial basis function for radius and quality modeling.

%Assume that there exists a slowly-time-varying scalar or vector field $\vx$ whose value needs to be estimated at a given number of discrete locations. To explore unknown spatially-distributed fields, autonomous agents are usually required to traverse the entire unknown environment while collecting sensory data to estimate a corresponding map (Figure~\ref{fig:ADMM-example-application}). We model the map as a set of locations of interests arranged on a regular grid, where at each location, we assume that there is a corresponding slowly-time-varying quantity. For our purposes, we are interested in finding paths that best reconstruct the field, i.e., that achieve the minimum uncertainty.

%We address the estimation problem using Kalman Filters (KFs)(cf. \cite{anderson2012optimal}), with a focus on the estimated covariance $P_j$, to calculate the map model uncertainty. We assume the robot provides field measurements at waypoints~$q$, with Gaussian radial basis function modeling radius and quality.
%\rtron{The previous paragraph can be written more compactly.}

@@

  \begin{definition} Let $\nu_\cF$ and $\nu_\cE$ be two unitary vectors ($\norm{\nu_\cF}=\norm{\nu_\cE}=1$). Define the normalized vector $u$ as
    \begin{equation}
      u'=\nu_\cF+\nu_\cE,\quad u =\frac{u'}{\norm{u'}}.
    \end{equation}
    The \emph{Householder rotation} $H(\nu_\cF,\nu_\cE)$ is defined as
    \begin{equation}\label{eq:H definition}
      H(\nu_\cF,\nu_\cE) = 2 u u\transpose-I.
    \end{equation}
  \end{definition}
  The main property of interest for our application is the fact that $H$ is a rotation mapping $\hat{\nu_\cF}$ to $\hat{\nu_\cE}$, as shown by the following.
  \begin{proposition}\label{prop:HProperty}
    The matrix $H$ has the following properties:
    \begin{enumerate}
    \item It is a rotation, i.e.
      \begin{enumerate}
      \item\label{it:orthonormality} $H\transpose H=I$;
      \item\label{it:determinant} $\det(H)=1$.
      \end{enumerate}
    \item\label{it:transformation} $\nu_\cE=H \nu_\cF$.
    \end{enumerate}
  \end{proposition}
  \begin{proof}
  	See Appendix \ref{proof:HProperty}.
  \end{proof}

  We compute the differential of $H$ implicitly using the relation \eqref{equ:dt_to_dx}. We will use the notation $\cross{v}:\real{3}\to\real{3\times 3}$ to denote the matrix representation of the cross product with the vector $v$, i.e.,
  \begin{equation}
    \bmat{v_1\\v_2\\v_3}\mapsto \bmat{0 & -v_3 &v_2\\v_3 & 0 & -v_1\\-v_2 & v_1 & 0},
  \end{equation}
  such that $[v]_\times w=v\times w$ for any $w\in\real{3}$. %One can verify by direct computation the following property:
  %\begin{equation}\label{eq:asymmetric to cross}
  %  wv\transpose-vw\transpose=\cross{\cross{v}w}.
  %\end{equation}
  \begin{proposition}\label{prop:Hderivitive}
    Let $\nu_\cF(t)$ represent a parametric curve. Then we have
    \begin{equation}
      \dot{H}=H\cross{-2M\dot{\nu}_{\cF}}
    \end{equation}
    where the matrix $M\in\real{3\times3}$ is given by
    \begin{equation}
      M=[u]_{\times}  \frac{ \left( I - u u\transpose \right) \left( I- \nu_\cF \nu_\cF\transpose \right)} {\norm{u'} \norm{\nu_\cF}}.
    \end{equation}
  \end{proposition}
  \begin{proof}
  	See Appendix \ref{proof:Hderivitive}.
  \end{proof}

@@

We now provide the functions $D(\vq)$, the sets $\cZ$, and the corresponding projection operators $\Pi_\cZ$ for traditional path planning constraints (\Cref{sec:velocity-constraint}-\Cref{sec:waypoint-constraint}), co-observation security constraints (\Cref{sec:co-observation-constraint}), and reachability constraints (\Cref{sec:??}). The latter are based on the definition of \emph{ellipsoidal reacability regions} (\Cref{sec:reachability}).
\subsection{Velocity constraints}\label{sec:velocity-constraint}
  The movement of each agent is constrained by a maximum distance in any direction over a single discrete time step. We then define the function $D(\vq)$ to return the velocity vectors for the $i$-th agent at time step $j$; its result then needs to be constrained to a sphere.

  \begin{constraint}[Velocity constraint]
\begin{align}
     D_{ij}(\vq) &= q_{ij}-q_{i(j-1)}, \quad j\in\{1,\ldots, T\}\\
\label{eq:velocity_constraint}
   \sZ_{ij} &=\bigl\{z \mid \norm{z} \leq v_{max}\bigr\}\\
   \Pi_{\sZ_{ij}}(z)&=\begin{cases}
   v_{\max}\frac{z}{\norm{z}} \textrm{ if } \norm{z}>v_{\max}\\
   z\textrm{ otherwise}
   \end{cases}
\end{align}
\end{constraint}

\subsection{Convex polygonal obstacles collision constraints} \label{sec:obstacle-constraint}

@@

\begin{constraint}[Obstacle constraint]
\begin{align}
     	D(q_{ij}) &= \max_{k=1,\dots,l} (d_{k}(q_{ij},\cP))\\
\label{eq:obstacle_constraint}
  \sZ &= \{z \mid z \geq 0 \},\\
   \Pi_\sZ(z) & = \begin{cases}
   0 & \textrm{if} \quad z < 0\\
   z  & \textrm{otherwise}
   \end{cases} \label{eq:zoneConstraintProjection}
\end{align}
\end{constraint}

The motivating idea for \eqref{eq:obstacle_constraint} and \eqref{eq:zoneConstraintProjection} is to identify waypoints within a specific region and project them to the closest boundary. Non-convex obstacles can be handled by the union of (possibly overlapping) convex obstacles. By setting the normal vector pointing toward the inside of the region, this method can also serve to constrain the robots within the designated workspace region.

\subsection{Waypoints with flexible deadlines}\label{sec:waypoint-constraint}
It is common for security related constraints to require robot reach a specific location at specific time. We assume robot $i$ is required to reach a given point $p$ within a radius of $d_{max}$ at some time instant $j$ belonging to a given time window $[t_1,t_2]$. The constraint function $D(\vq)$ to return the smallest distance from the point $p$ to waypoint on the trajectory restricted to the relevant time window. 
\begin{constraint}[Waypoint constraint]
\begin{align}
     	D(\vq) &= \min_{j\in\{t_1,\dots,t_2-1\}}\bigl(\dist(p,\overrightarrow{q_{ij}q_{i(j+1)}})\bigr)\\
\label{eq:waypoint_constraint}
  \sZ &= \{z \mid z < z_{max} \},\\
   \Pi_\sZ(z) & = \min(z, d_{max}).
\end{align}
\end{constraint}

where $\dist(p,\overrightarrow{q_{ij}q_{i(j+1)}})$ returns the distance between the fixed point $p$ and the segment $\overrightarrow{q_{ij}q_{i(j+1)}}$. Note that this function returns the smallest distance between $(p,q_{ij})$ and $(p,q_{i(j+1)})$ if the projection of the point $p$ does not lie on the line segment~$\overrightarrow{q_{ij}q_{i(j+1)}}$; as a consequence, this constraint does not need to be satisfied exactly at one of the points on the discretized trajectory, but it can also be satisfied ``en route'' on the segment between them.
\rtron{Give example application here, and include application of ADMM without the security constraints, but highlighting a path that would break the security}
\zyang{plot new simpler example, add collision avoidance}

%\subsection{Secure Multi-robot path planning using co-observation and reachability constraints}
\subsection{Co-observation schedule constraint}\label{sec:co-observation-constraint}
The co-observation constraints is used to guarantee that, at time required by the schedule, two robots should get close enough with each other to observe each other's behavior. This constraint is modeled as a relative distance constraint between two robots at some time instant $j$ (i.e., to require two agents see each other within a certain radius to inspect each other, or to exchange data). The location of the co-observation location should make sure that the ellipsoidal reachability region between each co-observation, considered in later sections \ref{sec:reachability}, has an empty intersection with all forbidden regions.
 We write the function for the constraint as:
\begin{constraint}[Co-observation constraint]
\begin{align}
D(\vq) &= \overrightarrow{q_{aj}q_{bj}}, \label{eq:waypoint_constraint}\\
  \sZ &= \{z \mid z \leq d_{max} \},\\
   \Pi_\sZ(z) & = \begin{cases}
d_{max}\frac{z}{\norm{z}} &\text{if } \norm{z} > d_{max},\\
z	& otherwise.
\end{cases}
\end{align}
\end{constraint}

where $a,b$ are the indices of the pair of agents required for a mutual inspection. The locations $q_{aj}$ and $q_{bj}$ where the co-observation performed are computed as part of the optimization.

@@

The reachability ellipsoid is an over-approximation of the exact reachability region; the difference between the two is due to the discretization of the trajectory, and the fact that $\cE$ does not consider the presence of obstacles.

\subsubsection{Transformation to canonical coordinates}\label{sec:rotation2Standard}
To simplify the problem, we employ a canonical rigid body transformation that repositions the ellipse $\cE$ from the global frame $\cF$ to a canonical frame $\cF_\cE$. In this new frame, the ellipsoid's center aligns with the origin, and its foci align with the first axis of $\cF_\cE$. We perform this transformation in a differentiable manner, with respect to the two foci serving as waypoints for a robot.

For the convenience of derivation, we define the coordinate transformation from $\cF$ to $\cF_\cE$ using a rotation $R^\cF_\cE$ and a translation $o^\cF_\cE$, which, to simplify the notation, we simply refer to as $R$ and $o$, respectively. The transformation of a point from the frame $\cF_\cE$ to the frame $\cF$ and its inverse are given by the formulation:
  \begin{equation}
    \Fframe{q}=R\Eframe{q}+o,\label{eq:E to F transform}\quad
    \Eframe{q}=R\transpose(\Fframe{q}-o).
  \end{equation}

  We define $\nu_\cF$ and $\nu_\cE$ to represent the $x$-axis unitary vector of $\cF_\cE$ in the frames $\cF$ and $\cF_\cE$, respectively. Formally:
  \begin{align}
    \nu'_\cF&=q_2-q_1, & \nu_\cF &= \frac{\nu'_\cF}{\norm{\nu_\cF}}, & \nu_\cE&=[1,0,0]^T;
  \end{align}
  see Fig.\ref{fig:EllipseConstraintExample} for an illustration. Note that $\nu_\cE$ is constant while, for the sake of clarity, $\nu_{\cF}$ depends on $q_1,q_2$. We then define the rotation $R$ using a \emph{Householder rotation}, while from \eqref{eq:E to F transform} we see that $o$ represents the center of $\cE(q_1,q_2)$ expressed in $\cF$, i.e.:
  \begin{align} 
    R&=H(\nu_\cF(q_1,q_2),\nu_\cE(q_1,q_2)), &o&=\frac{1}{2}(q_1+q_2).
  \end{align}
  In the following, we will consider $H$ to be a function of $q_1,q_2$ directly, i.e. $H(q_1,q_2)$. 
  The ellipse $\cE$ expressed in $\cF_\cE$ is given by $\Eframe{\cE}=\{\Eframe{q}\in\real{m}:d(q^\cE_1,\Eframe{q})+d(\Eframe{q},q^\cE_2)<2a\}$, where the coordinates of the two foci $q^\cE_1,q^\cE_2$ in $\cF_\cE$ are
  \begin{align}
    q^\cE_1=\bmat{c & 0 & 0}\transpose, && q^\cE_2=\bmat{-c & 0 & 0}\transpose,
  \end{align}
  with $c=\frac{\norm{q_2-q_1}}{2}$.
\begin{definition}
  The ellipsoid $\cE$ in the canonical frame can be described as the zero level set of the quadratic function
  \begin{equation}\label{equ:standard-ellipse}
    \Eframe{E}(\Eframe{q}) = \Eframe{q}\transpose Q \Eframe{q} - 1
  \end{equation}
  where
  \begin{equation}
    Q = \diag(a^{-2},b^{-2},b^{-2}),
  \end{equation}
  and $b = \sqrt{a^2-c^2}$.
  The ellipse parameters $a$,$b$ represent the lengths of the major axes.
\end{definition}

@@

\subsection{Point-ellipsoid reachability constraint}\label{sec:ellipsoid-point}
As shown in Fig.\ref{fig:Ellipse-to-point}, we consider a forbidden region in the shape of a single point $q_{avoid}$. The goal is to design the trajectory $q(t)$ such that $q_{avoid}\notin\mathcal{E}(q_1,q_2)$. This constraint can be written as,

\begin{constraint}[Point-ellipsoid reachability constraint]
\begin{align}
D(\vq) &=  \begin{cases}
      \pi_{p\cE}(q) - q_{avoid} & q_{avoid} \in \cE, \\
      {0} & \textrm{otherwise}.
    \end{cases} \label{eq:point_ellipsoid_constraint}\\
  \sZ &= \{q\in\mathbb{R}^{nm}:\norm{D_p(q)} = 0 \},\\
   \Pi_\sZ(z) & = 0, 
\end{align}
\end{constraint}

where $\pi_{p\cE}(q_{avoid};q_1,q_2,a) = q_p$ is a projection function that returns a projected point $q_p$ of $q_{avoid}$ outside the ellipse, i.e., as the solution to
\begin{equation}\label{eq:ellipse-point-projection}
\pi_{p\cE}=\argmin_{q_p\in\cE^c} \quad \norm{q_{avoid}-q_p}^2 
\end{equation}
where $\cE^{c}$ is the set complement of the region $\cE$.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\linewidth, trim = 1.5cm 2.5cm 1.5cm 3cm, clip]{Ellipse2point}
\caption{Point-ellipsoid constraint, a point $q_{avoid}$ inside ellipsoid is projected to the areas outside the ellipsoid $f_{p}$.}
\label{fig:Ellipse-to-point}
\end{center}
\end{figure}

  For cases where $q_{avoid} \notin \cE(q(t_2),q(t_1),r)$, $\pi_{p\cE}(q_{avoid}) = q_{avoid}$. And for cases where $q_{avoid} \in \cE(q_1,q_2,r)$, $D(q)$ needs to be projected to the boundary of the ellipse.

  \zyang{can this be treated as proposition}
  \begin{proposition} \label{prop:ellipse-project-canonical}
  In canonical frame $\cF_{\cE}$, the projection from a point $q^{\cE}_{avoid}\in \cE$ inside the ellipse to the boundary of $\cE$:
  \begin{equation}
  	q^\cE_{p} = {(I+sQ)}\inverse q^\cE_{avoid} = S q^\cE_{avoid}
  \end{equation}
  %where $S=(I+sQ)\inverse$.Using the fact that since $q^\cE_p$ is a point on the ellipse, 
  where $s$ can be solved as the root of the equation obtained by substituting \eqref{eq:qp from S} in $\Eframe E(\Eframe q)$:
  \begin{equation}
    0=F(s)={q^\cE_p}\transpose Q q^\cE_p-1={q^\cE_{avoid}}\transpose Q'(s) q^\cE_{avoid} -1,
  \end{equation}
  where
  \begin{equation}
    Q'(s) =S\transpose Q S = \diag\left(\frac{a^{2}}{(s+a^{2})^2}, \frac{b^{2}}{(s+b^{2})^2},\frac{b^{2}}{(s+b^{2})^2}\right)
  \end{equation}
 Detailed methods for computing $s$ can be found in \cite{eberly}. 
  \end{proposition}
  \begin{proof}
  Proof can be found in Appendix~\Cref{proof:ellipse-project-canonical}
  \end{proof}

@@

  %The point to project, $q_{avoid}$, can be likewise expressed in $\cF_\cE$ as $\Eframe{q}_{avoid} =  H (\Fframe q_{avoid}-o)$. We now turn our attention to the problem of projecting $\Eframe{q}_{avoid}$ on the zero level set of $E^\cE$ (i.e., the reachability ellipsoid in the canonical frame). The derivations below are loosely inspired by \cite{eberly}.

  %Let $q^\cE_p$ be the point on the surface of the ellipsoid, i.e., $\Eframe E(q^\cE_p) = 0$, corresponding to the projection of the point $q^\cE_{avoid}$. Using Lagrange multipliers applied to the constrained optimization problem \eqref{eq:ellipse-point-projection} (after transforming it in the canonical frame), one can show that the vector from a point to its projection, $q^\cE_{avoid}- q^\cE_p$, must be collinear with the gradient of $\Eframe \cE$, i.e.
 % \begin{equation}
 %   q^\cE_p- q^\cE_{avoid} = s \partial_q \Eframe E(q^\cE_p)\transpose = s Q q^\cE_p
  %\end{equation}
  %for some scale $s\in\real{}$;
 % thus $q^\cE_p$ can be written as:

  Then the point-to-ellipse projection function in $\cF$ can be represented as:
  \begin{multline}\label{equ:Point2EllipseProjection}
    \pi_{p\cE}(q)= R^{-1}(q(t_1),q(t_2))q^\cE_p+o \\
    = R^{-1}(q(t_1),q(t_2))Sq^\cE_{avoid}+o\\
    = R^{-1}S R ( q_{avoid}- o)+o
  \end{multline}

  In our derivations, we consider only the 3-D case ($m=3$); for the 2-D case, let $P=\bmat{I & 0}\in\real{2\times 3}$: then $\pi_{p\cE}^{\textrm{2D}}=P\pi_{p\cE}^{\textrm{3D}}(P\transpose q_{avoid}; P\transpose q_1, P\transpose q_2,a)$.
